Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-04-13 src/tex/ms.tex
Minor text change

% What's new?
In this study, we use a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
Compared to other studies, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
We consider 71 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
We discuss single-event results in detail, and identify degeneracies between birefringence and spin effects, in addition to the already known correlations with source orientation and distance.
We use the results from individual events to place a collective population constraint on the strength of \ac{GW} amplitude birefringence from GWTC-3.

@@

We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
This measurement is signficantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
Future catalogs with more events will shed light on this feature.

>>>>

% What's new?
In this study, we use a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
Compared to other studies, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.
We discuss single-event results in detail, and identify degeneracies between birefringence and spin effects, in addition to the already known correlations with source orientation and distance.
We use the results from individual events to place a collective population constraint on the strength of \ac{GW} amplitude birefringence from GWTC-3.

@@

We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
This measurement is significantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
Future catalogs with more events will shed light on this feature.
